### 2018-07-22 Chapters/Chapter1.tex
style changes

\begin{section}{Free particle in a periodic potential}

The one particle part of the Hamiltonian can be written as:

@@

\begin{equation}
\psi_{in}(\textbf{r}) = \frac{1}{\sqrt{M}}\sum_{\textbf{k}\in BZ} e^{-i\textbf{k}\textbf{r}_i} \psi_{n\textbf{k}}(\textbf{r})
\end{equation}

Where $M$ is the number of lattice sites as defined in \ref{AP1A}. Using \ref{Bloch1} we see that $\psi_{in}(\textbf{r}) = \psi_{0n}(\textbf{r}-\textbf{r}_i) \equiv \psi_{n}(\textbf{r}-\textbf{r}_i)$, so we only need to define one Wannier function for each band and the others are obtained by translations.

@@

The single particle Hamiltonian including spin orbit interaction is $\hat{H}^{(1)}(\textbf{r}) = -\frac{\hbar^2}{2m}\nabla^2_{\textbf{r}} + U(\textbf{r}) + \lambda\boldsymbol{\sigma}\boldsymbol{\nabla}V(\textbf{r}) \times \textbf{p}$, where $\lambda = \frac{\hbar^2}{4m^2c^2}$. Including interaction, the Hamiltonian for a system of $N_e$ electrons has the form:

Where $v$ represents the electron-electron interaction. We use the Wannier functions $\Phi_j$ as a basis and restrict ourselves to a fixed band so we don't write the band index. Let $\hat{c}_{j \sigma}^\dagger$ and $\hat{c}_{j \sigma}$ create and annihilate a particle in the state $\Phi_j$ and spin $\sigma$. In this basis, the second quantized form of \ref{Ham1} is:

@@

Where $t_{ij} = \bra{i} -\frac{\hbar^2}{2m}\nabla^2_{\textbf{r}} + U(\textbf{r}) \ket{j}$ and $\boldsymbol{\Delta}_{ij} = \bra{i} \lambda \boldsymbol{\nabla}V(\textbf{r}) \times \textbf{p} \ket{j}$. And where we assume $\hat{H}^{(1)}$ and $\hat{v}$ to be spin independent. Also we can impose that $t_{ii} = t(0) = 0$. For a system of strongly correlated electrons we can assume that:

@@

\begin{section}{Heisenberg antiferromagnet from Hubbard model}

@@

Where $\boldsymbol{D}_{ij} = -\frac{4it}{U}\boldsymbol{\Delta}_{ij}$.

>>>>

In this chapter we will start with a Hamiltonian

where $v$ is the electron-electron interaction and where $\hat{H}^{(1)}$ is the single particle Hamiltonian.

\begin{equation}
\hat{H}^{(1)}(\textbf{r}) = -\frac{\hbar^2}{2m}\nabla^2_{\textbf{r}} + U(\textbf{r}) + \lambda\boldsymbol{\sigma}\boldsymbol{\nabla}V(\textbf{r}) \times \textbf{p}
\end{equation}

Where $\lambda = \frac{\hbar^2}{4m^2c^2}$ and $U(\bs{r})$ is a periodic potential. In this Chapter we will derive an effective Hamiltonian using time independent perturbation theory.

\begin{section}{Single particle in a periodic potential}

First we will define a suitable basis for the single particle wavefunctions. To do so, we will consider the single particle Hamiltonian without the spin orbit interaction term:

@@

\begin{equation}
\phi_{in}(\textbf{r}) = \frac{1}{\sqrt{M}}\sum_{\textbf{k}\in BZ} e^{-i\textbf{k}\textbf{r}_i} \psi_{n\textbf{k}}(\textbf{r})
\end{equation}

Where $M$ is the number of lattice sites as defined in \ref{AP1A}. Using \ref{Bloch1} we see that $\phi_{in}(\textbf{r}) = \phi_{0n}(\textbf{r}-\textbf{r}_i) \equiv \phi_{n}(\textbf{r}-\textbf{r}_i)$, so we only need to define one Wannier function for each band and the others are obtained by translations. We will restrict ourselves to a fixed band, so we will drop the band index and write $\phi_i(\bs{r})$ or $\ket{i}$.

@@

In this section we will derive the second quantized form of \ref{Ham1} using the Wannier function basis. Let $\hat{c}_{j \sigma}^\dagger$ and $\hat{c}_{j \sigma}$ create and annihilate a particle in the state $\phi_j$ and spin $\sigma$. In this basis, the second quantized form of \ref{Ham1} is:

@@

Where $t_{ij} = \bra{i} -\frac{\hbar^2}{2m}\nabla^2_{\textbf{r}} + U(\textbf{r}) \ket{j}$ and $\boldsymbol{\Delta}_{ij} = \bra{i} \lambda \boldsymbol{\nabla}V(\textbf{r}) \times \textbf{p} \ket{j}$. Notice that $\bs{\Delta}_{ij} = - \bs{\Delta}_{ji}$. We assumed $\hat{v}$ to be spin independent. Also we can impose that $t_{ii} = t(0) = 0$. For a system of strongly correlated electrons we can assume that:

@@

\begin{section}{Effective Hamiltonian for half-filling system}

@@

Where $\boldsymbol{D}_{ij} = -\frac{4it}{U}\boldsymbol{\Delta}_{ij}$. The total effective Hamiltonian is:

\begin{equation}
\hat{H}_{eff} = \sum_{\langle i,j \rangle} J\boldsymbol{S}_i\boldsymbol{S}_j + \boldsymbol{D}_{ij} \boldsymbol{S}_i \times \boldsymbol{S}_j
\end{equation}
